\documentclass[a4paper,11pt]{book}

\usepackage{amsmath}
\usepackage{amssymb}
\usepackage[amssymb]{SIunits}
\usepackage{graphicx}
\usepackage[table]{xcolor}
\usepackage{enumitem}
\usepackage{fancyhdr}
\usepackage[footnotesize,bf]{caption}
\usepackage{sidecap}
% \usepackage{longtable}
% \usepackage{multirow}
% \usepackage{booktabs}
\usepackage[pdfborder={0 0 0}]{hyperref}
% \usepackage{float}

% \usepackage{titlesec}
% \titleformat{\chapter}[hang]{\bf\huge}{\thechapter}{0.5em}{}
% \titlespacing{\chapter}{0pt}{0pt}{50pt}

\usepackage{natbib}
\setlength{\bibsep}{1pt}
\renewcommand{\bibfont}{\footnotesize}

\hoffset -1in
\voffset -1in

\topmargin      10mm
\headheight     10mm
\headsep        10mm
\oddsidemargin  25mm
\textwidth     160mm
\textheight    242mm

%\title{\bf DNS/TMP User Guide}
\title{{\bf AT}\textcolor{black!50}{Lab} Documentation}
\author{Turbulence and Boundary Layers Group $|$ University of Hamburg}

\setcounter{tocdepth}{2}
\setlength{\parindent}{0mm}

\fancyhead[LE,LO]{\footnotesize\leftmark}
\fancyhead[RE,RO]{}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\def\pst{\partial_t}
\def\psk{\partial_k}

\def\Re{\mathrm{Re}}
\def\Pr{\mathrm{Pr}}
\def\Sc{\mathrm{Sc}}
\def\Fr{\mathrm{Fr}}
\def\Ro{\mathrm{Ro}}
% \def\Le{\mathrm{Le}}
% \def\Ma{\mathrm{Ma}}
% \def\Da{\mathrm{Da}}

\def\j{\mathbf{j}}
\def\e{\mathbf{e}}

\def\bg{_\mathrm{bg}}
\def\ss{_\mathrm{s}}
\def\sc{_\mathrm{c}}

\newcommand{\pt}[2] {\partial_{#2}{#1}}
\newcommand{\ptp}[2]{\partial_{#2}\left(#1\right)}
% \newcommand{\ptp}[1]{\partial_t\left(#1\right)}
% \newcommand{\pjp}[1]{\partial_j\left(#1\right)}
\newcommand{\dvp}[1]{\nabla\negmedspace\cdot\negthinspace\left(#1\right)}
\newcommand{\dv} [1]{\nabla\negmedspace\cdot\negthinspace#1}
\newcommand{\avr}[1]{\overline{#1}}
\newcommand{\avf}[1]{\widetilde{#1}}


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{document}

\frontmatter
\pagestyle{empty}
\maketitle
\tableofcontents

\setlength{\parskip}{0.5\baselineskip}
%\include{preface}

\chapter*{Preface}
\addcontentsline{toc}{chapter}{Preface}
\sloppy

ATLab---an acronym for Atmospheric Turbulence Laboratory---is a set of tools whose aim is {\bf to efficiently solve and analyze a particular set of governing equations with a controlled accuracy}. The accuracy can be controlled in different ways: comparing with analytical solutions, including linear stability analysis; grid convergence studies; balance of transport equations, like integral turbulent kinetic energy or local values at specific relevant locations (e.g., at the wall). Resolution can be measured by the ratio between the grid spacing $\Delta x$ and the relevant small scales, like the Kolmogorov scale $\eta$ or the thickness of the diffusion sub-layers next to the wall. For the compact schemes used here, typical values are $\Delta x/\eta\simeq 1-2$; larger values can lead to numerical instability because of the aliasing generated by the non-linear terms. Note that these schemes are non-monotone, but typical out-of-bounds deviations of conserved scalars are below $10^{-6}-10^{-8}$ relative to the mean variations, and this error is therefore negligibly small compared to the typical error associated with the statistical convergence, of the order of $1-5$\%. The statistical convergence can be estimated by varying the sample size of the data set, e.g. varying the domain size along the statistically homogeneous directions. 

The efficiency can be measured in different ways but, ultimate, it should be related with the computational time needed to understand a particular problem with a given accuracy, and so the importance of the controlled accuracy. 

Making the code user-friendly comes after the previous two main priorities: controlled accuracy and efficiency.

Last, the main documentation is the code itself and the examples. This document is only a short introduction to the equations and the tools. The code is continuously under development, so this document is continuously incomplete.

\mainmatter
\pagestyle{fancy}

\part{User Guide}
\include{equations}
% \include{bcs}
% \include{code}
% \include{grid}
\include{postprocessing}

\part{Technical Guide}
% \include{numerical}
\include{memory}
% \include{decomposition}
% \include{scaling}
\include{profiling}

\backmatter
\bibliographystyle{plainnat}
\bibliography{atlab-doc.bib}

\end{document}
